% Restructured book: Chapter 1 imported from Chap 1_ Motivation and Objectives.docx.
\chapter{Introduction}

\section{Motivation}

Interior layout design is a practical need in Vietnam's urban housing context. In 2024, the Ministry of Construction reported about 125,545 successful apartment and individual-house transactions \cite{Vietnam2025InformationHousingReal}, showing a large housing market where many households may face furnishing, renovation, or reorganization decisions. In residential spaces, users often need to decide how furniture should be arranged to make a room functional, comfortable, and visually coherent. This includes choosing suitable positions for beds, wardrobes, sofas, desks, storage units, and media furniture while considering the room shape, available area, openings, and daily circulation. These decisions affect cost, circulation, and daily usability. If the initial layout ignores room boundaries, openings, or furniture dimensions, users may need avoidable revisions before purchasing furniture or implementing the arrangement.

To obtain a satisfying home, customers often hire architects or interior designers, but professional design services require additional cost and time. In Vietnam, these services typically cost around 150,000–250,000 VND/m\textsuperscript{2}, and the design process often takes 10–20 days depending on project complexity and revisions \cite{NoiThatMansion2026ApartmentDesignPrice,Hometalk2026ApartmentInteriorDesignPrice}. Therefore, a lightweight tool can support two practical user groups: users who want to design the room by themselves, and users who may hire a professional later but still want to explore initial layout ideas first.

To support initial layout design exploration, the first challenge is to help users create an early furniture layout by themselves. The target users include people arranging their own bedrooms or living rooms, apartment occupants comparing furniture options, and small accommodation owners who need a quick spatial concept before professional refinement. For these users, a useful layout must do more than place objects inside a room. It must respect the room boundary, furniture dimensions, doors, windows, collision avoidance, and basic circulation. It should also remain editable because early requirements are often incomplete and may change after the user sees the first design.

The second challenge is to turn the selected layout into a realistic and understandable visual representation. A top-down layout is useful for checking positions and distances, but it does not fully show how the room may feel in practice. From a 2D plan alone, users may still find it difficult to imagine materials, color harmony, lighting, depth, and overall atmosphere. A rendered perspective image helps them evaluate whether the selected layout is visually acceptable before they continue revising the design or purchasing furniture.

These two problems are closely connected. Layout generation provides spatial control and editability, while image rendering provides visual feedback. If the rendered image is generated without respecting the layout, it may look attractive but become difficult to use as a design reference. If the system only provides an editable layout without realistic visualization, the user may still struggle to assess the design. Therefore, this thesis focuses on two connected problems in early-stage interior design: generating an editable and geometrically valid initial furniture layout for non-professional users, and producing realistic visualization that remains consistent with the selected layout.

\section{Objectives and Scope of the Thesis}

An early-stage interior design system should help users turn an initial room idea into a result that they can inspect, revise, and use for further decisions. This is still difficult because existing solutions usually stop at only one side of the process.  Image generation tools can quickly produce attractive interior pictures, but the pictures may not follow the actual room shape or remain editable as layouts. Research systems for room arrangement can suggest furniture placement, but the result still needs to be checked against room boundaries, doors, windows, collisions, and circulation. These limitations show that the main gap is not only image quality or layout suggestion, but the lack of a connected workflow from user requirements to an editable layout and a layout-consistent realistic image.

Based on this gap, this thesis defines two main objectives. The first objective is to build a system that can generate an initial furniture layout from an ordinary-language user brief, explicit room geometry, fixed openings, and a furniture inventory. The generated layout must be stored as structured room-and-object data so that it can be checked for basic spatial constraints, displayed in 2D/3D, and manually refined by the user through direct editing. The second objective is to generate realistic perspective images from the selected layout. These images should help users understand the visual direction of the room while preserving the main room structure, camera view, openings, and furniture arrangement defined by the editable layout. In addition, these two objectives are integrated into a web application so that users can create rooms, generate layouts, preview designs, render images, and edit results within one continuous workflow.

The scope of this thesis is limited to living rooms and bedrooms. These room types are selected because they are common in apartments and contain representative layout constraints. A living room usually involves seating, tables, media furniture, storage, and circulation relationships. A bedroom usually involves bed placement, wardrobe access, bedside furniture, lighting, and optional work or entertainment areas. The system takes as input a natural-language design brief, a room boundary, fixed openings, and a furniture inventory. In this thesis, the inventory is a database of furniture records with category, dimensions, style and material metadata, and optional 3D asset files. The expected outputs are an editable 2D layout, a 3D preview, a generated render image, and controlled image-editing results.

The thesis does not aim to replace architects or professional interior designers. It also does not address construction-level tasks such as structural design, electrical and plumbing systems, technical lighting design, ventilation, cost estimation, or legal construction drawings. The proposed system is intended as an early-stage design assistant. Its purpose is to help users explore possible layouts, revise them interactively, and obtain visual feedback before moving to more detailed design decisions.

\section{Proposed Approach}

To achieve these objectives, the thesis proposes RoomFlow, an AI-assisted room design system that combines language understanding, geometric validation, editable layout data, and layout-based visualization. The key idea is that RoomFlow should not generate a final image directly from a text prompt. Instead, it first builds a structured room layout that can be checked, edited, and previewed. This layout then becomes the basis for 3D preview, perspective rendering, and later image editing. In this way, the proposed solution keeps spatial control in the editable layout stage while using image generation mainly to improve visual realism and user understanding.

In the layout design stage, the system takes the room type, room boundary, fixed openings, furniture inventory, and user brief as input. A multi-agent planning pipeline interprets the user’s intent, identifies required furniture, and infers high-level spatial preferences. The final placement is then checked by a geometric solver rather than relying on the language model alone. This separation allows the system to validate containment, collisions, opening clearance, and circulation before returning an editable layout that users can review and modify.

The 3D preview acts as a bridge between the editable layout and the realistic image generation stage. After a layout is selected, RoomFlow reconstructs it as a 3D scene using Three.js\footnote{\url{https://threejs.org/} is a JavaScript library for rendering interactive 3D graphics in the browser.}. The user can inspect the room from a perspective view, adjust the camera, and choose the view that best represents the design. This selected view becomes the visual reference for generating the realistic image.

In the realistic image generation and editing stage, the system uses the selected layout, 3D snapshot, camera view, visible furniture, and preservation constraints to guide the image model. The goal is to improve materials, lighting, depth, and atmosphere while preserving the room shell, openings, camera framing, and major furniture arrangement. The same stage also supports localized image editing after the initial render is generated. Users can request changes such as adding, removing, replacing, or recoloring selected elements, while unrelated parts of the room are expected to remain stable.

Overall, the proposed solution treats layout data as the central representation of the design. The language model helps convert user intent into design requirements, the validation module ensures that the layout remains spatially usable, the 3D preview helps users inspect the arrangement, and the image generation module provides realistic visual feedback. By connecting these components, the system supports the two main problems of this thesis: creating editable initial layouts and producing layout-consistent visualization for early-stage interior design.


\section{Thesis Organization}

The remainder of this thesis is organized into eight chapters.

Chapter 2 surveys existing approaches to AI-assisted interior design and analyzes the system requirements. The chapter reviews representative solutions related to interior layout design, realistic interior image generation, localized image editing, and integrated design workflows. From their limitations, the chapter identifies the main requirements of RoomFlow and specifies the key use cases, functional requirements, and non-functional requirements.

Chapter 3 establishes the theoretical background for layout design and realistic image generation. The chapter introduces LLM-assisted planning, structured agents, indoor scene synthesis, asset-grounded layouts, layout-conditioned image generation and editing, evaluation metrics, and the web technologies used in the system.

Chapter 4 presents the automatic interior layout design method. The chapter describes the overall processing flow, input representation, semantic planning module, room modeling module, object program construction, geometric solver, and preparation of editable layout outputs.

Chapter 5 presents the realistic image generation and editing method. The chapter explains how the selected layout is converted into visual guidance for perspective rendering, how visible-object information and preservation constraints are used to preserve layout consistency, and how localized image editing is controlled after the initial render is generated.

Chapter 6 describes the implementation of RoomFlow as a web application. The chapter presents the software architecture, package design, user interface design, application pipeline, 3D preview and rendering workflow, database design, main implementation tools, implemented functions, and deployment configuration.

Chapter 7 evaluates automatic interior layout design. The chapter introduces the layout benchmark dataset and metrics, explains the metrics through representative cases, compares RoomFlow with I-Design and FlairGPT, and presents the layout ablation study.

Chapter 8 evaluates realistic image generation and localized image editing. The chapter introduces the rendering and editing benchmark datasets, defines the preservation and edit-success metrics, compares six rendering variants, evaluates localized image editing, and discusses preset fidelity, runtime, and preservation trade-offs.

Chapter 9 concludes the thesis and outlines future work. The chapter summarizes the main results, discusses the remaining limitations, and suggests possible directions for improving layout design, rendering consistency, localized editing, and user interaction.
